\section{Conclusion \& Future Work}\label{sec:conclusion&future-works}
\subsection{Conclusion}
This paper presented a large-scale ablation study of shortterm electricity 
price forecasting in Denmark, training 1,052
models across eight architectures, 13 feature constellations,
eight forecasting horizons, and two prediction targets. Rather
than optimising for peak performance, the study examined how
default-configured models respond to systematic variation in
training conditions, with the practical aim of identifying which
data sources contribute meaningfully to predictive accuracy.

Two primary findings emerge from the experimental results.
First, tree-based models, in particular CatBoost, achieved
the lowest \gls{mae} overall and demonstrated the strongest
performance at short horizons. Neural network architectures
underperformed at the 24h horizon under default training
settings. However, as horizons expanded this gap began
minimizing, suggesting different architectural trade-offs apply at different 
timescales. Second, feature contribution analysis
showed that historical price information was consistently the
most impactful single feature group. When comparing the
general performance of models in experiments with the pricing
group to ones without it, when targeting absolute price, no
experiment that included the price group performed worse than
any experiment without it. Results when targeting the price
delta are similar, where comparing an experiment to itself with
and without the price group, the experiment including the price
group almost always performed better. By contrast, adding
lagged grid features beyond the immediate present produced
a decrease in performance, indicating that the consumption
patterns of previous days do not assist models in predicting
the price in the future.

The supplementary experiments validated several methodological choices made in 
the main study. Loss function
alignment produced substantial improvements for poorly-performing feature sets 
but only marginal gains for the best-performing ones, suggesting that loss 
alignment is a useful
safety net against suboptimal feature selection. Optuna-based
walk-forward hyperparameter tuning displayed substantial improvements of 
22-33\% across all tree-based models and both
prediction targets. This suggests that default hyperparameters
are sub-optimal and that targeted hyperparameter tuning offers a
reliable path to improvement. The substitution of \gls{dmi} weather
data with Midas-provided weather, under fair date-matched
comparison, showed no practical difference in model performance across 
architectures and targets. Adding Midas data to
already existing datasets showed varied results, depending on
target, supporting the broader observation that larger quantities
of data do not directly result in better performance. The result
suggests that the choice of weather provider may be flexible,
and that comprehensive data preparation by external providers
does not yield a measurable advantage over the per-station
aggregation methodology used in this paper.

Regarding the central question of how much data is required,
the results support a pragmatic conclusion: a forecasting system
can achieve competitive accuracy with weather, time, and
historical price features alone, without the need for additional
data gathering and preprocessing. The marginal accuracy gains
from comprehensive feature collection may not, depending on
the particular use-case, justify the corresponding increase in
data acquisition and maintenance complexity. This finding has
direct practical implications for any organisation considering
deployment of such a system in the Danish electricity market.

The limitations of this study should be acknowledged.
All models were trained with default hyperparameters under
fixed walk-forward validation settings, and the supplementary
experiments suggest that targeted tuning could shift the relative
performance of specific architectures. The neural network
experiments were necessarily constrained in coverage due to
computational costs, limiting the strength of architecture-family
comparisons at longer horizons. Finally, all findings are specific
to the Danish day-ahead market, particularly for energy region
DK1. The performed cross-region comparison experiments
between DK1 and DK2 suggests reasonable robustness, but
extension to other electricity markets would require further
validation.

\subsection{Future work}\label{sec:futurework}
To further investigate this topic, one could start with the
following subjects:
\paragraph{Analysis of neighbouring countries} Due to the
interconnectedness of the European energy market, being able
to accurately predict the production and consumption of neighbouring countries 
would likely improve the accuracy of models
for individual countries. However, doing so poses the problem
of gathering data from multiple different governments and
institutions with varying reporting standards and requirements.
The creation of a compound energy prediction model for the
whole of Europe would likely be the most accurate, but is left
for future, ambitious, researchers.
\paragraph{Specialised neural networks} As stated within section II, 
specialized neural networks have seen promising results
in various prediction scenarios. A deeper analysis into how
such models may be adjusted for the Danish energy market
may yield better performance than displayed in this paper.
\paragraph{Including alternative electricity production prices}
During this paper, conditions that impact the price and
availability of fossil fuels for energy production was not
considered. One can imagine that factoring in the price of
coal, oil, and natural gas would likely impact the price of
energy as a whole.
\paragraph{Investigation differing models on an individual weather
station basis} The supplementary experiment looking into the
generalizability of prediction models into Danish energy region
DK2 suggested generalizability, but results varied depending
on target and architecture. This suggests that a full analysis
of DK2 may be justified, and that further specialization of
models, such as models targeted towards individual weather
stations close to renewable energy sources, may yield even
more accurate results.
\paragraph{Mid-term prediction to confirm neural net applicability}
As horizons expanded, all models seemed to experience a
drop-off in performance. To further investigate this, one could
expand the horizon, to see when performance may shift, perhaps
leading to the neural network-based models gaining the upper
hand.