\section{Experiments} \label{sec:experiments}
This section contains details of all the performed experiments, as well as 
other considerations related thereto.
\subsection{Experiment suite}\label{sec:experimentSuite}
All models were trained over data from an 11 year time-frame, from January 2015 
to mid-March 2026. They were also trained in a walk-forward expanding-window 
with 
an initial training window of two years, then a test window of 30 days, as well 
as a step size of 30 days. As previously mentioned, data-groups were 
constructed in 
order to create different data-scenarios for training. These groups consisted 
of the following categories: 
\begin{itemize}
	\item \textbf{Weather}: All weather-related features. 
	\item \textbf{WeatherLags}: T-Minus 24 hour data of the Weather group. 
	\item \textbf{Grid}: All data related to \gls{pc}, aside from the foreign 
	energy exchange features.
	\item \textbf{GridLags}: T-Minus 24 hour data of the Grid group. 
	\item \textbf{GridExchange}: Only the data related to foreign energy 
	exchange. 
	\item \textbf{GridExchangeLags}: T-Minus 24 hour data of the GridExchange 
	group. 
	\item \textbf{Price}: Columns containing the absolute price and price 
	delta. These are not included in the training data, but are instead 
	individually used as prediction targets depending on the type of experiment 
	being run, either absolute price or price delta. 
	\item \textbf{PriceLags}: T-Minus 24, 48, and 
	168 hour data of both absolute 
	price and price delta. 
	\item \textbf{Time}: Data related to time, such as hour, day of the week, 
	day of the month, day of the year, as well as sine and cosine 
	representations.
	\item \textbf{AllFeatures}: Simply a collection of all previous groupings.
\end{itemize}

Based on these groups, we created 13 base experiments, generally increasing 
information sequentially: 
\begin{enumerate}
	\item Weather, Time
	\item Weather, WeatherLags, Time
	\item Weather, PriceLags, Time
	\item Weather, WeatherLags, PriceLags, Time
	\item Weather, Grid, GridExchange, Time
	\item Weather, WeatherLags, Grid, 
	GridExchange, Time
	\item Weather, Grid, GridExchange, PriceLags, 
	Time
	\item Weather, WeatherLags, Grid, 
	GridExchange, PriceLags, Time
	\item Weather, Grid, GridExchange, GridLags, 
	GridExchangeLags, Time
	\item Weather, WeatherLags, Grid, 
	GridExchange, GridLags,	GridExchangeLags, 
	Time
	\item Weather, Grid, GridExchange, GridLags, 
	GridExchangeLags, PriceLags, Time
	\item Weather, WeatherLags, Grid, 
	GridExchange, GridLags, GridExchangeLags, 
	PriceLags, Time
	\item AllFeatures
\end{enumerate}
These 13 experiments were run at 0, 24, 48, 72, 96, 120, 144, and 168-hour 
horizons for both, targets absolute price and price delta. Note that price 
delta was 
not used as a target for horizon 0 hour, as the delta between the current and 
target would always be 0. The 24 hour horizon was run for all eight models, 
with the subsequent horizons being run for all tree-based models, but only for 
select neural network-based models, due to their training-time constraints. 
This resulted in $4 \times 13 \times (7+8) = 780$ tree-based models, where 4 is 
the number of tree-based model types, 13 is the number of feature set 
configurations, and $(7+8)$ represents the 8 absolute price horizons and 7 
price delta horizons respectively (horizon 0h is excluded from the delta target 
as it is always zero). For the neural network-based models and AutoGluon, an 
equivalent set of 4 model types were evaluated, yielding $4 \times (13 \times 2 
+ 3 \times 7 \times 2) = 272$ models, where the first term covers all 13 
feature sets at the 24h horizon for both targets, the second term covers the 3 
selected feature sets across the remaining 7 horizons and both targets, and the 
factor of 2 in both terms represents the two prediction targets. The reduced 
number of neural network-based experiments was determined by their 
computational constraints, with the 3 selected feature sets identified as the 
best, mean, and worst performers based on their 24h horizon results. This 
results in a total of $780 + 272 = 1052$ models trained, with
a smaller number of additional, supplementary experiments
discussed in \autoref{sec:expSupExp}.

\subsection{Evaluation criteria}\label{sec:evaluationCriteria}
While there are many different criteria one could choose to focus on when 
comparing the performance of models, the one chosen for this paper is 
\gls{mae}. The equation for calculating this metric is shown in 
\autoref{eq:mae}.
\begin{equation}
	\text{MAE} = \frac{1}{n} \sum_{t=1}^{n} |y_t - \hat{y}_t|
	\label{eq:mae}
\end{equation}
where $y_t$ and $\hat{y}_t$ are the ground truth target value at 
prediction time 
and the predicted value at prediction time respectively.\\ 
In this context, this metric represents the average amount of EUR/MWh 
the model was off. This metric was partly chosen due to the fact that it 
is comparable across various model architectures and target types. 
Furthermore, it is expressed in the same unit as the target, making it 
easily interpretable. Lastly, in a market with frequent price spikes, 
both positive and negative, it does not penalise 
these extremes as 
harshly as other criteria. Other considered metrics are discussed in 
\cref{app:A:equations}, and all of the values, both for the chosen and 
non-chosen metrics are available online in the GitHub repository. 
Additionally, a naive baseline, that simply reuses yesterday's price, 
calculated as follows: $\hat{y}_t = y_{t-24}$ was computed across all 
horizons to establish a non-ML reference point. 

\subsection{Experimental setup}
All model training, validation, and testing were conducted locally on a machine 
running Microsoft Windows 11 Home, equipped with a 13th Gen Intel Core 
i7-13650HX processor and 16 GB of RAM, utilising 
CPU computation exclusively. 
The exclusion of the GPU was evaluated on the basis that the libraries used for 
training and validation of the neural network-based models were optimised for 
the CPU. This means that while training time was a somewhat limiting factor for 
the neural network-based models in this paper, this aspect may be mitigated 
with different training approaches and hardware, however, the performance 
metrics, aside from training time, should be unaffected.

\subsection{Training, testing \& validation}\label{sec:split_protocol}
Model evaluation in this paper is performed using walk-forward validation with 
an expanding training window. The procedure works as follows: An initial 
training window covering January 2015 through December 2016 is used to train
each model for the first time. The model is then evaluated on
the immediately following 30-day test window. After evaluation,
the training window expands to include those 30 days, and
the test window slides forward to the next 30-day period.
This process repeats until the full dataset is exhausted. The
expanding window design ensures that all available historical
data is utilised at each step, while the fixed 30-day test window
provides a consistent evaluation unit across all iterations. A
figure illustrating this process is shown in \autoref{fig:walkForward}. 
\begin{figure}[t]
	\centering
	\resizebox{\linewidth}{!}{%
		\begin{tikzpicture}[
			x=1cm, y=0.8cm,
			train/.style={rectangle, fill=darkgray, minimum height=0.5cm},
			test/.style={rectangle, fill=orange, minimum height=0.5cm},
			font=\sffamily
			]
			
			% Axis
			\draw[->, thick, >=stealth] (0, 0.5) -- (9.5, 0.5) node[above left, 
			yshift=2pt] {\textbf{Present}};
			\node[above right, yshift=2pt] at (0, 0.5) {\textbf{Time}};
			
			% Pass 1
			\node[left] at (0, -0.5) {\textbf{Pass 1}};
			\node[train, minimum width=3.5cm, anchor=west] at (0.2, -0.5) {};
			\node[test, minimum width=1.5cm, anchor=west] at (3.8, -0.5) {};
			
			% Pass 2
			\node[left] at (0, -1.2) {\textbf{Pass 2}};
			\node[train, minimum width=4.2cm, anchor=west] at (0.2, -1.2) {};
			\node[test, minimum width=1.5cm, anchor=west] at (4.5, -1.2) {};
			
			% Pass 3
			\node[left] at (0, -1.9) {\textbf{Pass 3}};
			\node[train, minimum width=4.9cm, anchor=west] at (0.2, -1.9) {};
			\node[test, minimum width=1.5cm, anchor=west] at (5.2, -1.9) {};
			
			% Pass 4
			\node[left] at (0, -2.6) {\textbf{Pass 4}};
			\node[train, minimum width=5.6cm, anchor=west] at (0.2, -2.6) {};
			\node[test, minimum width=1.5cm, anchor=west] at (5.9, -2.6) {};
			
			% Pass 5
			\node[left] at (0, -3.3) {\textbf{Pass 5}};
			\node[train, minimum width=6.3cm, anchor=west] at (0.2, -3.3) {};
			\node[test, minimum width=1.5cm, anchor=west] at (6.6, -3.3) {};
			
			% Available Historical Time Series Base Bar
			\node[rectangle, fill=gray!80, minimum width=9cm, minimum 
			height=0.4cm, anchor=west] at (0.2, -4.2) 
			{\color{white}\textbf{Available Historical Time Series}};
			
			% Legend
			\node[train, minimum width=0.8cm, anchor=west, 
			label=right:Training] at (2.5, -5.2) {};
			\node[test, minimum width=0.8cm, anchor=west, 
			label=right:Forecasting] at (5.0, -5.2) {};
			
		\end{tikzpicture}%
	}
	\caption{Figure visualising the method of training and validating
		a model through utilisation of walk-forward expanding-window
		validation.}
	\label{fig:walkForward}
	\vspace{-4mm}
\end{figure}
While this approach of essentially retraining the model 12 times per
year for every year in the training data, leading to about 100
validation steps per model, is therefore computationally
expensive, it is the generally preferred method of training when
dealing with time-series data ~\cite{tashmanWalkForward}. This approach was 
chosen over standard k-fold cross-validation, which randomly partitions
data into folds without regard for temporal ordering. Applying
k-fold to time-series data would allow a model to be trained on
future observations and tested on past ones, constituting a form
of data leakage that would produce optimistically biased results.
Walk-forward validation avoids this by strictly preserving the
temporal ordering of time-series data. The model is always
trained exclusively on historical data and evaluated on data it
has never seen. 

\subsection{Data leakage prevention}
The perceptive reader may have noticed that the master matrix dataset retains 
the raw spot price and all target columns to serve as a shared foundation for 
all experiments. However, feeding this matrix directly to a model would 
constitute severe data leakage, as the model could trivially learn to copy the 
target value rather than forecast it. To prevent this, a Target Shield is 
applied at load time before each training run, which drops all leaky columns 
from the feature set. Specifically, the active prediction target is removed 
from the features, the raw spot price is excluded except when included as a 
lagged historical feature, and the concurrent price delta columns are removed. 

\subsection{Feature importance}
After training various constellations and horizons of many
of the models, we analysed the feature importance through
permutation importance on the best performing models of
each architecture at each horizon. Permutation importance is
an inspection technique applied by systematically dropping
each column individually and measuring the impact on the
evaluation criteria. For this paper, it was deemed that if
the dropping of a column resulted in even a drop of 0,01
EUR/MWh, then it would be allowed to stay. This method
is an extensively rigorous, though computationally expensive,
approach to feature importance. The results of these pruned
models will be discussed in \autoref{sec:results}.

\subsection{Supplementary experiments}\label{sec:expSupExp}
To address generalisability and the limitations that come
with running the models on their default settings i.e. with no
hyperparameter tuning and with library-default loss functions,
optimised for \gls{mse}, five supplementary experiments were conducted:
\paragraph{Loss function alignment} The best-performing neural
network model, as identified from the 24h baseline results, was
retrained using \gls{mae} loss instead of \gls{mse} to evaluate whether
loss-metric alignment would improve performance.
\paragraph{Hyperparameter tuning} Optuna-based 
hyperparameter optimisation 
was applied to tree models on the best, mean, and worst performing feature sets 
at 24h for both targets, with 30 trials per configuration, to gauge the 
improvement such tuning may yield.
\paragraph{AutoGluon quality escalation} AutoGluon was allowed to train using 
the medium\_quality preset and a 180
second time limit. This was done on the best feature set at the
24h-horizon, and was compared against the baseline fastest
preset, with a six second time limit, that was used on the
ordinary experiments
\paragraph{Midas Energy data} Internally used weather features
from Midas Energy were merged with the master matrix
and tested on all models at the 24h-horizon and on the tree-based models at all 
horizons, to compare against the raw \gls{dmi}
station-level aggregation. Additionally, it was also investigated
how substituting the general weather data used in the paper
for the Midas data would impact predictive performance
\paragraph{DK2 generalisation check} The best 
tree-based and
neural network-based models were trained on DK2 data at a
24h-horizon across best, mean, and worst feature sets to verify
that DK1 findings generalise regionally.\\
The results for all of these supplementary experiments may be
found in \autoref{sec:resultsSupExp}, along with a comparison to a naive
baseline, to verify whether a naive approach may compete with
machine learning architectures for this task.
