\section{Introduction} \label{sec:introduction}
In a world of increasing amounts of electric vehicles, data centres, and other 
large-scale electricity-hungry industries, electricity grids are beginning to 
feel the strain~\cite{ppm-gridImpact}, contributing to an increase in 
electricity prices~\cite{gridStrainIncreasingGraph}. When faced with such 
increased prices, it is beneficial, from a monetary point of view, to schedule 
electricity intensive tasks during periods where prices are low. However, to do 
this, one must know when such low-price periods occur. In Denmark, the price of 
electricity is, in simple terms, fixed for the next 24-hour period at noon 
daily\footnote{This is a simplification for ease of reading. For actual energy 
trading details one may visit \cite{nordpoolTrading}. It is also further 
explained in \autoref{sec:prereq}}. 

This is often sufficient for household electricity use, as one can typically 
schedule dishwashers, washing machines, and electric vehicle charging to happen 
during the lowest point of that period. This can be done either through direct 
scheduling on individual appliances or smart devices, or through smart home 
management systems such as Home Assistant~\cite{homeAssistant}. Industrial 
needs for energy can involve larger and more time-sensitive electricity 
demand~\cite{industrialHouseholdEnergyConsumption}, and may therefore be more 
expensive. As an example, one might imagine a near-future where fleets of 
electric trucks need their batteries charged before delivery. In a scenario 
like this, it may be beneficial for such companies, to be able to forecast 
electricity prices further into the future, and plan their electricity 
consumption accordingly, to minimize charging costs. \\

When attempting to predict electricity prices, many different aspects can be 
considered~\cite{forecastInput}. Among these, the ones of particular interest 
to this paper are: 
\begin{itemize}
	\item Weather
	\item Electricity production details
	\item Electricity consumption details
	\item Historical- or T-minus data\footnote{Data from an earlier point in 
	time before the current time, e.g. 24 hours before.}
	\item Time of year/month/week/day
	\item Electricity exchange with foreign countries
\end{itemize}

One factor that influences electricity prices is the weather, due to its direct 
impact on both electricity production~\cite{productionPatterns} and 
consumption~\cite{consumptionPatterns}. In a country that has invested 
extensively in wind-energy, such as Denmark, the weather directly impacts 
electricity production, as more than $80\%$ is produced by renewable 
sources~\cite{energiStyrelsenEnergiStatestik2022}. Steady winds, and to a 
lesser extent, clear skies therefore affect energy production quite heavily in 
Denmark, as necessities for efficient wind and solar energy production. When it 
comes to consumption, human consumption patterns also change depending on the 
weather~\cite{consumptionPatterns}. When it is cold, people turn on the heat to 
warm themselves, and when it is hot, they turn on air conditioning, or fans, to 
stay cool. Further, when it is cold, such as in the winter months, it is 
typically also darker, necessitating the need to use lights more frequently. 

These factors suggest that one can use the weather to predict electricity 
\gls{pc} patterns, which implies that weather data may carry useful information 
for electricity price forecasting. A substantial body of prior work has 
examined short-term electricity price forecasting using statistical, 
machine-learning, and deep-learning 
approaches~\cite{energyPricePredictionMLNeighbors,energyPricePredictionUK,energyPricePredictionNovelDeepLearning},
 showing various results in different forecasting scenarios. 
However, much of this prior work shares common limitations. Results are 
typically reported for a single model trained on a fixed feature set, often 
utilizing highly specialized and computationally expensive deep-learning 
architectures~\cite{energyPricePredictionNovelDeepLearning}. While these 
approaches achieve strong benchmarks, they may be difficult and costly for 
industry practitioners to deploy, maintain, and retrain in production 
environments. This obscures a fundamental practical question: what data 
actually impacts the performance of models, and can standard, out-of-the-box 
algorithms perform comparably without the overhead? Someone wishing to deploy a 
forecasting system is therefore left without clear guidance on whether 
investing in comprehensive data collection and complex model development is 
worthwhile, or whether a simpler feature set utilizing a general-purpose model 
would suffice. Further, while related studies exist, the Danish setting 
combines several factors that are not jointly emphasized in most of the 
literature reviewed here, such as its heavy investment into wind-energy 
~\cite{energiStyrelsenEnergiStatestik2022}, its division of the energy grid 
into two distinct bidding zones, and its heavy energy-exchange with its 
neighbouring countries. \\

This paper addresses these gaps through a large-scale ablation study, 
systematically varying the feature composition, forecasting horizon, and 
prediction target across 1,052 trained models. Rather than reporting the best 
achievable performance under ideal, heavily engineered conditions, the goal is 
to answer a practically motivated question: how much data do you actually need? 
This is not merely an academic exercise, but a question of commercial 
viability. An energy trading company deploying a real-time forecasting system 
must weigh the cost of acquiring and maintaining each data source against the 
marginal accuracy it may provide. Understanding which feature categories 
contribute meaningfully, and at which horizons they stop being useful, has 
direct commercial value.\\

To ground this research in real-world application, this project was conducted 
in partnership with Midas Energy~\cite{midasEnergy}. Midas Energy currently 
provides short-term electricity production and consumption forecasts to energy 
trading companies, but is actively looking to expand into direct price 
forecasting. Throughout this study, they contributed essential commercial 
context, data acquisition support, and benchmarking frameworks, which will be 
discussed where relevant.\\

To investigate this dynamic across a broad spectrum of architectures, eight 
different types of general-purpose machine learning algorithms and frameworks 
are evaluated: 
\begin{multicols}{2}
	\begin{enumerate}
		\item RandomForest\cite{randomForestRegressor}
		\item XGBoost\cite{xgboost}
		\item LightGBM\cite{lightgbm}
		\item CatBoost\cite{catboost}
		\item LSTM\cite{lstm}
		\item GRU\cite{gru}
		\item Transformer\cite{transformer}
		\item AutoGluon\cite{autogluon}
	\end{enumerate}
\end{multicols}
Model training is based on weather data gathered from the 
\gls{dmi}, as well as electricity \gls{pc} data from \gls{en}. To 
comprehensively evaluate the chosen forecasting approaches, the experiments are 
structured around three distinct prediction settings. First, the models are 
trained on various data constellations, ranging from purely weather-related 
data to a more comprehensive dataset of weather, grid, and historical price 
data. This isolates the impact of each specific data category. Second, the 
forecasting horizon is expanded sequentially to evaluate how the accuracy of 
each algorithm deteriorates as it predicts further into the future. Finally, 
the approaches are trained using two distinct prediction targets: predicting 
the absolute electricity price and predicting the price delta. For the purpose 
of this paper, the price delta refers to the change in price from the current 
hour to the forecasted hour. This dual-target approach is motivated by a desire 
to anticipate model overfitting and to investigate whether the algorithms are 
more adept at capturing relative price fluctuations rather than absolute 
prices. The overall purpose of this ablation study is to uncover how different 
machine learning frameworks react when given identical training conditions 
under these varying settings.\\

Producing a meaningful ablation study in this setting required substantial 
preprocessing effort. The raw data sources used for this paper span nearly 11 
years, and consist of 
station-level weather measurements, energy production and consumption records, 
and price data across two format changes. This data was not designed for 
machine learning consumption. Constructing a consistent, temporally aligned 
training dataset required geographic filtering of weather stations, cross-year 
column homogenisation, resolution standardisation, and imputation of 
structurally missing data. This preprocessing pipeline is a necessary 
foundation for the validity of the experimental results, and is described in 
detail in \autoref{sec:data_preprocessing} to enable reproducibility.\\

The experimental results yield two primary findings. First, standard tree-based 
models, particularly CatBoost, dominate short-term horizons, outperforming more 
complex neural networks in their default configurations. Second, while weather
and time data establish a baseline, historical price data is the single most 
critical feature group; expanding the dataset to include lagged grid 
consumption features yields diminishing or even negative returns. Supplementary 
experiments further demonstrate that targeted hyperparameter tuning provides 
substantial accuracy gains. Additionally, substituting the custom-made, 
aggregated, station-level weather data with professionally processed weather 
data provides no significant predictive advantage. Ultimately, this study 
demonstrates that industry practitioners can achieve respectively competitive 
forecasting accuracy using out-of-the-box algorithms and a constrained feature 
set, avoiding the significant overhead of specialized deep learning 
architectures and excessive data acquisition.\\
